---
title: "Untitled"
author: "Zachary M. Smith"
date: "June 13, 2018"
output: html_document
---

# Visual Exploration

## All Reference

The following figure shows the distribution of reference data with two plots colored by the reference data correctly identified as "EGF" (blue) and those incorrectly identified as "PVP" (orange). __Figure a__ is a scatter plot depicting the relationship between Chessie BIBI score and percent impervious surface. __Figure b__ groups the data by the rating categories ("EGF" and "PVP") and presents the data as box-and-whisker plots overlaid with jittered points.
```{r, fig.width=8, fig.height=8}

point.plot <- ggplot(ref.imp, aes(final_score, imp_value, color = ref_cat)) +
  geom_point(alpha = 0.25) +
  scale_color_manual(values = c("EGF" = "#56B4E9",
                               "PVP" = "#E69F00"),
                     name = "Rating") +
  geom_hline(aes(yintercept = 2), linetype = "dashed") +
  geom_label(aes(-5, 2, label = "2%"), color = "black") +
    geom_hline(aes(yintercept = 5), linetype = "dashed") +
  geom_label(aes(-5, 5, label = "5%"), color = "black") +
  geom_hline(aes(yintercept = 10), linetype = "dashed") +
  geom_label(aes(-5, 10, label = "10%"), color = "black") +
  # facet_wrap(~ref_cat) +
  ylab("Impervious Cover (%)") +
  xlab("BIBI Score")

box.plot <- ggplot(ref.imp, aes(ref_cat, imp_value, color = ref_cat)) +
  
  geom_jitter(alpha = 0.25) +
    scale_color_manual(values = c("EGF" = "#56B4E9",
                               "PVP" = "#E69F00"),
                     name = "Rating") +
  geom_boxplot(outlier.alpha = 0, alpha = 0, color = "black") +
  geom_hline(aes(yintercept = 2), linetype = "dashed") +
  geom_label(aes(0.5, 2, label = "2%"), color = "black") +
    geom_hline(aes(yintercept = 5), linetype = "dashed") +
  geom_label(aes(0.5, 5, label = "5%"), color = "black") +
  geom_hline(aes(yintercept = 10), linetype = "dashed") +
  geom_label(aes(0.5, 10, label = "10%"), color = "black") +
  ylab("Impervious Cover (%)") +
  xlab("BIBI Score")

cowplot::plot_grid(point.plot, box.plot, ncol = 1, labels = "auto")
```

## Bioregion Reference

Scatter plots by bioregion. __Warning:__ The y-axes are not standardized to make it easier to view the full distribution of data relative to each bioregion. The dashed horizontal line representing 2% impervious cover should enable the reader to quickly orient themselves.
```{r, fig.width=8, fig.height=10}
ref.imp %>% 
  left_join(bio.fam.df, by = c("event_id", "sample_number", "final_score")) %>% 
ggplot(aes(final_score, imp_value, color = ref_cat)) +
  geom_point() +
  scale_color_manual(values = c("EGF" = "#56B4E9",
                               "PVP" = "#E69F00"),
                     name = "Rating") +
  geom_hline(aes(yintercept = 2), linetype = "dashed") +
  geom_label(aes(-5, 2, label = "2%"), color = "black") +
    geom_hline(aes(yintercept = 5), linetype = "dashed") +
  geom_label(aes(-5, 5, label = "5%"), color = "black") +
  geom_hline(aes(yintercept = 10), linetype = "dashed") +
  geom_label(aes(-5, 10, label = "10%"), color = "black") +
  xlim(-10, 100) +
  facet_wrap(~spatial, ncol = 2, scales = "free_y") +
  ylab("Impervious Cover (%)") +
  xlab("BIBI Score")
```

Violin plots by bioregion. Violin plots are similar to box-and-whisker plots but they provide more information about the distribution of data; the wider the polygon the more data points present. The median was added to each violin plot as a black horizontal line. __Warning:__ The y-axes are not standardized to make it easier to view the full distribution of data relative to each bioregion. The dashed horizontal line representing 2% impervious cover should enable the reader to quickly orient themselves.
```{r, fig.width=8, fig.height=10}
ref.imp %>% 
  left_join(bio.fam.df, by = c("event_id", "sample_number", "final_score")) %>% 
ggplot(aes(ref_cat, imp_value, color = ref_cat, fill = ref_cat)) +
  geom_violin(draw_quantiles = c(0.5), color = "black") +
  # scale_color_manual(values = c("EGF" = "#56B4E9",
  #                              "PVP" = "#E69F00"),
  #                    name = "Rating") +
  scale_fill_manual(values = c("EGF" = "#56B4E9",
                               "PVP" = "#E69F00"),
                     name = "Rating") +
  geom_hline(aes(yintercept = 2), linetype = "dashed") +
  geom_label(aes(0.6, 2, label = "2%"), color = "black", fill = "white", alpha = 0.5) +
  facet_wrap(~spatial, ncol = 2, scales = "free") +
  ylab("Impervious Cover (%)") +
  xlab("BIBI Score")
```
