---
title: "Untitled"
author: "Zachary M. Smith"
date: "June 13, 2018"
output: html_document
---

# Preprocessing

Load necessary packages. __toolbox__ is a custom package for data preprocessing. The function `toolbox::prep_df()` will be used for imported data to standardize column names and character columns by converting all characters to lowercase, removing leading/trailing white space, removing all non-alphanumeric characters and replacing them with underscores ("_").
```{r, echo=FALSE}
suppressPackageStartupMessages(
  library(tidyverse)
)
library(toolbox)
```

## Import Data

Import family-level bioregion index scores and ratings.
```{r}
score.path <- "H:/Projects/Chessie_BIBI/report/FINAL_May25_2017/2017_Data/Scores_Ratings"
bio.fam.path <- "Bioregion/BIOREGION_FAMILY_All_Event_Ratings_2017-06-29.csv"

bio.fam.df <- data.table::fread(file.path(score.path, bio.fam.path)) %>%
  toolbox::prep_df() %>%
  mutate(date = as.Date(date, "%m_%d_%Y"),
         category = if_else(category == "sev", "deg", category))

```

Import Chessie BIBI stations assigned NHDPlus 2 Medium Resolution catchment ID (`featureid`).
```{r}
catchment.df <- data.table::fread(file.path(rprojroot::find_rstudio_root_file(),
                                            "data/assigned_catchments_oct16.csv")) %>%
  toolbox::prep_df()
```

Join the catchment information with the 2017 Chessie BIBI scores and ratings.
```{r}
bio.fam.df <- left_join(bio.fam.df, catchment.df,
                        by = c("station_id"))
rm(catchment.df)
```

This is a check that all rows now have a catchment assigned. Any rows without a catchment will appear in a table below the code chunk. The absence of a table indicates that all rows have an assigned catchment.
```{r}
test <- bio.fam.df %>%
  filter(is.na(featureid))
if(nrow(test) > 0) {
test %>% 
    kableExtra::kable() %>% 
  kableExtra::kable_styling(bootstrap_options = c("striped", "hover", "condensed"))
}
```

Import and join the Impervious Surfaces 2001, 2006, and 2011 StreamCat files.
```{r}
stream.cat.path <- "H:/GIS3/GIS5/New_GIS/streamcat/unzip"

imp.filename.vec <- c("ImperviousSurfaces2001_Region02.csv",
                      "ImperviousSurfaces2006_Region02.csv",
                      "ImperviousSurfaces2011_Region02.csv")

imp.df <- map(imp.filename.vec, function(file.i) {
  data.table::fread(file.path(stream.cat.path, file.i)) %>%
    toolbox::prep_df()
}) %>%
  reduce(full_join, by = c("comid", "catareasqkm",
                           "wsareasqkm", "catpctfull",
                           "wspctfull"))
```


## Assign Impervious Surfaces Data

Join the Impervious Surfaces data with the Chessie BIBI scores and ratings.
```{r}
imp.bibi.df <- left_join(bio.fam.df, imp.df, by = c("featureid" = "comid"))
```

Assign the Chessie BIBI data the correct Impervious Surfaces based on collection date. A range of ± 2.5 years around the Impervious Surfaces data collection years (2001, 2006, and 2011) was used to assign the Chessie BIBI data the most appropriate Impervious Surfaces values. __Therefore, any data collected prior to 1998-06-01 or after 2013-06-01 cannot be assigned an Impervious Surfaces value.__
```{r}
imp.bibi.df <- imp.bibi.df %>%
  mutate(
    # pctimp2001ws = if_else(date < as.Date("2003-06-01"), pctimp2001ws, as.numeric(NA)),
    pctimp2001ws = if_else(date >= as.Date("1998-06-01") & date < as.Date("2003-06-01"),
                           pctimp2001ws, as.numeric(NA)),
    pctimp2006ws = if_else(date >= as.Date("2003-06-01") & date < as.Date("2008-06-01"),
                           pctimp2006ws, as.numeric(NA)),
    pctimp2011ws = if_else(date >= as.Date("2008-06-01") & date < as.Date("2013-06-01"),
                           pctimp2011ws, as.numeric(NA))
    # pctimp2011ws = if_else(date >= as.Date("2008-06-01"), pctimp2011ws, as.numeric(NA))
  ) %>% 
  mutate(
    # pctimp2001cat = if_else(date < as.Date("2003-06-01"), pctimp2001cat, as.numeric(NA)),
    pctimp2001cat = if_else(date >= as.Date("1998-06-01") & date < as.Date("2003-06-01"),
                           pctimp2001cat, as.numeric(NA)),
    pctimp2006cat = if_else(date >= as.Date("2003-06-01") & date < as.Date("2008-06-01"),
                           pctimp2006cat, as.numeric(NA)),
    pctimp2011cat = if_else(date >= as.Date("2008-06-01") & date < as.Date("2013-06-01"),
                           pctimp2011cat, as.numeric(NA))
    # pctimp2011cat = if_else(date >= as.Date("2008-06-01"), pctimp2011cat, as.numeric(NA))
  )
```

## Reference Data

Subset the data to only include reference data.
```{r}
ref.df <- imp.bibi.df %>%
  filter(category == "ref") %>%
  mutate(ref_cat = if_else(final_score < ref_10, "PVP", "EGF"),
         ref_cat = factor(ref_cat, levels = c("PVP", "EGF"))) %>%
  arrange(final_score)
```

Make the three percent impervious surfaces columns a single column by transforming the data from a wide format to a long format. The column name `imp` represents the data source (e.g., `pctimp2006ws`) and the column name `imp_value` represents the percent impervious surfaces value. ~~All data collected prior to 1998-06-01 or after 2013-06-01 is maintained because impervious surfaces values could not be accurately assigned. For data collected between 1998-06-01 and 2013-06-01,~~ Rows with `NA` in the `imp_value` column indicate that the impervious surface year is not appropriate for the corresponding Chessie BIBI sample, and therefore were removed.
```{r}
ref.imp <- ref.df %>%
  select(event_id, date, sample_number, final_score, ref_cat,
         pctimp2001ws, pctimp2006ws, pctimp2011ws) %>%
  gather(imp, imp_value, pctimp2001ws:pctimp2011ws) %>%
  filter(!is.na(imp_value))
  # mutate(keep = case_when(
  #   date < as.Date("1998-06-01") | date >= as.Date("2013-06-01") ~ "keep",
  #   date >= as.Date("1998-06-01") & date < as.Date("2013-06-01") & !is.na(imp_value) ~ "keep",
  #   date >= as.Date("1998-06-01") & date < as.Date("2013-06-01") & is.na(imp_value) ~ "remove",
  #   TRUE ~ "ERROR"
  #   )) %>% 
  # filter(keep == "keep") %>% 
  # select(-keep)
```
