---
title: "Untitled"
author: "Zachary M. Smith"
date: "June 13, 2018"
output: html_document
---

# Reference Refinement

Apply the `cutpoints()` and `ce()` functions to the original data. The CE produced will be the baseline from which to compare the CE where reference sampling events with percent impervious surfaces upstream is greater than or equal to 2 are excluded.
```{r}
org.ce <- cutpoints(bio.fam.df) %>% 
  ce() %>% 
  mutate(org_ref_count = tp + fn,
         org_deg_count = tn +fp) %>% 
  select(spatial, ce, org_ref_count, org_deg_count)  %>% 
  rename(orginal_ce = ce)
```

This code chunk loops through and retains only impervious surface values less than 2, 5, or 10. The threshold used for a particular data subset is indicated in the colum `imp_thresh`. The `cutpoints()` and `ce()` functions  are applied to the data.
```{r}
imp.ce <- purrr::map_df(c(2, 5, 10), function(i) {
  ref.imp %>% 
    filter(imp_value >= i) %>% 
    anti_join(bio.fam.df, ., by = c("event_id", "sample_number", "final_score")) %>% 
  cutpoints() %>% 
  ce() %>% 
  mutate(new_ref_count = tp + fn,
         new_deg_count = tn + fp) %>% 
  select(spatial, ce, new_ref_count, new_deg_count) %>% 
  rename(new_ce = ce)  %>% 
    mutate(imp_thresh = paste("<", i))
})
```

Join data frame containing the original CE with the data frame containing the new CE values calculated after all reference sampling events with greater than or equal to 2% impervious surfaces upstream.
```{r}
comp.df <- full_join(org.ce, imp.ce, by = "spatial") %>% 
  mutate(change_ce = new_ce - orginal_ce,
         change_ref_count = new_ref_count - org_ref_count) %>% 
  mutate_if(is.numeric, funs(round(., 2))) %>% 
  arrange(desc(change_ce)) %>% 
  select(spatial, imp_thresh, org_ref_count, new_ref_count, change_ref_count,
         orginal_ce, new_ce, change_ce) 
```

Interestingly, the majority of bioregions had less than a 3% improvement in CE after removing reference sampling events with percent impervious surface values. MAC did have a 5% improvement in CE but the Reference count also dropped from 17 to 11. Removing sampling events from the reference condition if percent impervious surface is greater than or equal to 2 does lead to minor improvements but does not appear to be the main cause of the low scoring reference sampling events.
```{r}
comp.df  %>% 
  kableExtra::kable() %>% 
  kableExtra::kable_styling(bootstrap_options = c("striped", "hover", "condensed"))
```
## Reference 25th and 10th Percentiles

```{r}
org.quant <- bio.fam.df %>% 
  filter(category == "ref") %>% 
   group_by(spatial) %>% 
  summarize(org_quant25 = quantile(final_score, 0.25),
            org_quant10 = quantile(final_score, 0.10))
```

```{r}
imp.quant <- purrr::map_df(c(2, 5, 10), function(i) {
  ref.imp %>% 
    filter(imp_value >= i) %>% 
    anti_join(bio.fam.df, ., by = c("event_id", "sample_number", "final_score")) %>% 
  filter(category == "ref") %>% 
  group_by(spatial) %>% 
  summarize(new_quant25 = quantile(final_score, 0.25),
            new_quant10 = quantile(final_score, 0.10)) %>% 
    mutate(imp_thresh = paste("<", i))
})
```

```{r}
quant.df <- full_join(org.quant, imp.quant, by = "spatial") %>% 
  mutate(diff_25 = new_quant25 - org_quant25,
         diff_10 = new_quant10 - org_quant10) %>% 
  select(spatial, imp_thresh,
         org_quant25, new_quant25, diff_25,
         org_quant10, new_quant10, diff_10) %>% 
  arrange(spatial, imp_thresh)
```

```{r}
quant.df  %>% 
  kableExtra::kable() %>% 
  kableExtra::kable_styling(bootstrap_options = c("striped", "hover", "condensed"))
```


```{r, fig.width = 8, fig.height=25}
bio.fam.ref <- bio.fam.df %>% 
  filter(category == "ref") %>% 
  mutate(group = "original",
         quant_10 = quantile(final_score, 0.1))
i<- 2
test <- purrr::map_df(c(2, 5, 10), function(i) {
  test <- ref.imp %>% 
    filter(imp_value >= i) %>% 
    anti_join(bio.fam.ref, ., by = c("event_id", "sample_number", "final_score")) %>% 
  mutate(group = paste("<", i),
         quant_10 = quantile(final_score, 0.1))
}) %>% 
  bind_rows(bio.fam.ref) %>% 
  mutate(group = factor(group, levels = c("original", "< 10", "< 5", "< 2")))
  

ggplot(test, aes(group, final_score)) +
  geom_boxplot() +
  geom_errorbar(aes(ymin = quant_10, ymax = quant_10),
                width = 0.5,
                color = "red") +
  facet_wrap(~spatial, scale = "free", ncol = 1)
  
```

